Debiased Calibration for Generalized Two-Phase Sampling

### 2024-04-18 final/reports/prelim_paper.tex
Updated prelim_paper but need to fix part 1 theory.

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

@@

Combining information from several sources is an important practical problem. (CITEME)
We want to incorporate information from external data sources to reduce the bias
in our estimates or improve the estimator's efficiency. For many problems, the
additional information consist of summary statistics with standard errors. The
goal of this project is to incorporate external information with existing data 
to create more efficient estimators using calibration weighting.

To model this scenaro, we formulate the problem as a generalized two-phase
sample where the first phase sample consists of data from multiple sources. The
second phase sample contains our existing data. To motivate this setup, we
consider the following approach: first, we consider the classical two-phase
sampling setup where the second phase sample is a subset of the first phase
sample; then, we extend this setup to consider non-nested two-phase samples;
and finally, we consider the more general approach of having multiple sources as
the first phase sample.

@@

Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
an initial (Phase 1) sample of size $n$ is selected and $X_i$ is observed. Then
from the Phase 1 sample of elements, a (Phase 2) sample of size $r < n$ is
selected and $Y_i$ is observed. This is two-phase sampling (See 
\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
goal of two-phase sampling is to construct an estimator of $Y$ not
only using the observed information from the Phase 2 sample but also
incorporating the extra auxiliary information of $X$ from the Phase 1 sample, and
the challenge is doing this efficiently.

@@

$$ \pi_i(A_1) = \pi_{2i | 1} \pi_{1i}.$$

%\textcolor{red}{Also, the above equality is not true under two-phase sampling!)}
\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
  unnecessary. I have removed them. I have also modified the previous equation.
  I agree that the invariance condition does not hold for two-phase sampling,
  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
  because this is how conditional distributions work? (Also, I thought that this
  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
(1999), and Kim et al. (2006).) I would love to discuss this with you.}

@@

$$ \hat Y_{reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_2+ \sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)$$

where for $q_i = q(x_i)$ and is a function of $x_i$,
$$
\hat \beta_2 = \left(\sum_{i \in A_2} 
  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}. 
\footnote{
  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
  $\hat{\beta}_2$. Please check my book.}
  \textcolor{red}{Thanks for the check. I
checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
  more consistent with how I defined the two-phase regression estimator;
  however, I see your point about using $q_i$ to connect this estimator with the
superpopulation model.}}
  $$

@@

This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was extended by 
\cite{deville1992calibration} to consider loss functions other than just squared
loss. They generalized the loss function to minimize $\sum_i G(w_i, d_i)q_i$ for
weights $w_i$ and design-weights $d_i$ where $G(\cdot)$ is non-negative, strictly
convex function with respect to $w$, defined on an interval containing $d_i$,
with $g(w_i, d_i) = \partial G / \partial w$ continuous.
\footnote{The \cite{deville1992calibration} paper considers regression estimators
for a single phase setup, which we apply to our two-phase example. } This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others. The variance estimation is based on a linearization
that shows that minimizing the generalized loss function subject to the
calibration constraints is asymptotically equivalent to a regression estimator.

While this generalization is useful to an analyst who may want different
properties of their estimator from maximum entropy estimation rather than the
minimal squared loss, 
% TODO: I am not sure if this is correct -----
unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
estimator is not design consistent. 
\footnote{ \textcolor{blue}{I do not understand this part.} 
  \textcolor{red}{I changed the
notation so that it would not be the unused column space notation. This is from
Equation (11.17) of \cite{kim2024statistics}.}} 
\footnote{\textcolor{red}{I think that I am confused about the purpose of having
the assumption from Equation (11.17) in your sampling book. Could we discuss
this?}}
% --------------------------------------------

Furthermore, at a conceptual level, the regression estimator
has a nice feature that its two terms can be thought about as minimizing
variance and bias correction,

$$ \hat Y_{reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_2}{\pi_{1i}}}_{
  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_2)}_{
\text{Bias correction}}.$$

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. Instead, there is
a desire to get design consistency from the calibration term. In 
\cite{kwon2024debiased}, the authors show that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

The method of \cite{kwon2024debiased} requires known calibration
levels (no uncertainty) for the finite population. It does not handle the
two-phase setup where we need to estimate the finite population total of $x$
from the Phase 1 sample. Hence, we extend this method to have a valid variance 
estimator when including estimated Phase 1 weights.

\subsection*{Methodology}

@@

\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{x_i w_{2i}}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{x_i}{\pi_{1i}}
\end{equation}

@@

The first constraint is the existing calibration constraint and the second
ensures that design consistency is achieved. Here, 
$g(w) = \partial G / \partial w$. 
% TODO: Elaborate on this more.
The original method of \cite{kwon2024debiased} only considered having finite
population quantities on the right hand side of \ref{eq:calconst1}.
Let $z_i =
(x_i, g(\pi_{2i|1}^{-1})$. To solve this minimization problem, we use the convex
conjugate function $F(w) = -G(g^{-1}) + g^{-1}(w)w$, and instead find

$$ \hat \lambda = \argmin_{\lambda \in \Lambda_A} \sum_{i \in A_2} F(w_i) +
\lambda_1^T \sum_{i \in A_1} \frac{z_i}{\pi_{2i|1}}$$

for $\Lambda_A = \{\lambda: \lambda^T z_i \in g(\mathcal{V}) \text{ for all } i
\in A_2\}$.

\subsection*{Theoretical Results}

We prove two results in this section is the same spirit as \cite{kwon2024debiased}.
First, we state the assumptions of our analysis then we give the results.

\begin{itemize}
  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
    $\mathcal{V} \in \R$,
  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
    and $j$ in $A_2$, the Phase 2 sample and 
    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
    \infty,$$
    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
    $j$ in $A_1$, the Phase 1 sample and 
    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
    \infty,$$
  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
    = (\bf 0, 1)$.
\end{itemize}

\begin{theorem}[Design Consistency]
  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
  Equation~\ref{eq:calconst2} 
  exists and is unique with probability approaching 1. Furthermore, the
  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
  satisfies

  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$

  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
  \cite{kwon2024debiased}.

\end{theorem}

\begin{proof}
  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
  manner. There are two main modifications that need to be made. The first is
  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
  instead of a single $n$. The second is that we need to ensure that the result
  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
  consistent.

  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
  the first challenge is done. The second step largely holds from the existing
  proof of \cite{kwon2024debiased} with modifications such as

  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
  \lambda^T \sum_{i \in A_1} z_i^*$$

  ensuring that our extension holds.

  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
  detail, but I am unsure about the level of granularity that is required now. I
  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
  with the two-phase modification to show this result, but I don't know if I
  should yet.}
\end{proof}

We need the following additional assumptions to prove asymptotic normality.

% FIXME: Add assumptions
\textcolor{red}{I need to think about the assumptions that we need in two-phase
sampling.}
\begin{itemize}
  \item[(B1)] 
  \item[(B2)] The HT estimator is asymptotically normal under the sampling
    design in the sense that

\end{itemize}

\begin{theorem}[Asymptotic Normality]
% FIXME: Add result.
\textcolor{red}{I need to write out more but we should have something like this:
  $$V(\hat \theta_{DCE} = \Var(\hat \theta_{DC}) + \hat \gamma_{[1:3]}^T V(\bf
x)) \hat \gamma_{[1:3]} + \Cov(\hat \theta_{DC}, \bf x \hat \gamma_{[1:3}]).$$}
\end{theorem}

\textcolor{red}{Dr. Kim, I am not convinced that $\Cov(\hat \theta_{DC}, \bf x
\hat \gamma_{[1:3]}) = 0$. The reason that this covariance holds in
\cite{fuller2009sampling} Section 3.3.3 is that the estimating equation requires
$\sum_i x_i \hat e_i = 0$, which means that $\Cov(\hat e, x_i) = 0$. However, we
do not have this requirement. Instead, I see this more like a nested model where
we have a group level error that aggregates the individual level errors. The
problem with the current simulation that I cannot tell if this covariance should
be zero because the variance term $\hat \gamma_{[1:3]}'V(\bf x) \hat \gamma_{[1:3]}$
dominates the variance estimation. I will change the simulation parameters to
try to see if I can get a better test of this.}

\subsection*{Simulation Studies}

@@

\section*{Topic 2: Non-nested Two-Phase Sampling}

@@

In contrast to the classical two-phase sampling framework, non-nested two-phase sampling involves two independent surveys conducted on the same target population. The key distinction is that the two samples, denoted as $A_1$ and $A_2$, are drawn independently rather than sequentially.  Table \ref{table:11-1} presents the data structure for non-nested two-phase sampling.

In the non-nested two-phase sampling, 
a large probability sample $A_1$ is drawn from a finite population,  collecting  only the $\bx$ variable, and  a  smaller sample $A_2$ is drawn from the same population, providing information  on both the $y$ and $\bx$ variables.
It is assumed that the observed variable $x$ is comparable in the two surveys. \cite{renssen1997} formally addressed this non-nested two-phase sampling problem  and 
\cite{merkouris2004} extended the idea further to develop regression estimation combining information from multiple surveys. 
\cite{kimrao12} considered the non-nested two-phase sampling in the context of mass imputation combining two independent surveys at the population and domain levels.

\begin{table}[htb]
\caption{Data Structure for non-nested two-phase sampling}
\label{table:1}\par
\vskip .2cm
\centerline{\tabcolsep=3truept\begin{tabular}{|c|cc|}
\hline
Sample  & $X$ & $Y$ \\
\hline
$A_1$ & \checkmark &  \\
$A_2$ & \checkmark & \checkmark  \\
\hline
\end{tabular}}
\label{table:11-1}
\end{table}

>>>>

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

@@

Combining information from several sources is an important practical problem. (CITEME)
We want to incorporate information from external data sources to reduce the bias
in our estimates or improve the estimator's efficiency. For many problems, the
additional information consists of summary statistics with standard errors. The
goal of this project is to incorporate external information with existing data 
to create more efficient estimators using calibration weighting.

To model this scenaro, we formulate the problem as a generalized two-phase
sample where the first phase sample consists of data from multiple sources. The
second phase sample contains our existing data. To motivate this setup, we
consider the following approach: first, we consider the classical two-phase
sampling setup where the second phase sample is a subset of the first phase
sample; then, we extend this setup to consider non-nested two-phase samples;
and finally, we consider the more general approach of having multiple sources.

@@

Consider a finite population of size $N$ containing elements $(X_i, Y_i)$ where
an initial (Phase 1) sample of size $n_1$ is selected and $X_i$ is observed. Then
from the Phase 1 sample of elements, a (Phase 2) sample of size $n_2 < n_1$ is
selected and $Y_i$ is observed. This is two-phase sampling (See 
\cite{fuller2009sampling}, \cite{kim2024statistics} for general references.) The
goal of two-phase sampling is to construct an estimator of $\bar Y_N$ 
that uses both the observed information in the Phase 2 sample and also the extra
auxiliary information from $X$ in the Phase 1 sample.
The challenge is doing this efficiently.

@@

$$ \pi_i^{(2)}(A_1) = \pi_{2i | 1} \pi_{1i}.$$

%\textcolor{blue}{Also, the above equality is not true under two-phase sampling!)}
%\textcolor{red}{I agree that having the superscript $(1)$ and $(2)$ is
%  unnecessary. I have removed them. I have also modified the previous equation.
%  I agree that the invariance condition does not hold for two-phase sampling,
%  but if we consider $\pi_i$ as a function of $A_1$ then I believe this holds.
%  Is this correct? Should we not have $\pi_{i} = \pi_{1i} \pi_{2i|1}(A_1)$
%  because this is how conditional distributions work? (Also, I thought that this
%  is equivalent to the reverse sampling idea of Fey (1992), Shao and Steel
%(1999), and Kim et al. (2006).) I would love to discuss this with you.}

@@

$$ \hat Y_{reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} x_i \hat \beta_q+ \sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)$$

where for $q_i = q(x_i)$ and is a function of $x_i$,
$$
\hat \beta_q = \left(\sum_{i \in A_2} 
  \frac{x_i x_i'}{\pi_{1i} q_i}\right)^{-1} 
\sum_{i \in A_2} \frac{x_i y_i}{\pi_{1i} q_i}.$$ 
%\footnote{
%  \textcolor{blue}{Caleb, we do not have to use $\pi_{2i \mid 1}^{(2)}$ in computing
%  $\hat{\beta}_2$. Please check my book.}
%  \textcolor{red}{Thanks for the check. I
%checked it and I must have gotten $\pi_{1i}$ and $\pi_{2i}$ confused.}}
%\footnote{\textcolor{red}{I was initially thinking of using $\hat \beta_2$ without $q_i$ to be
%  more consistent with how I defined the two-phase regression estimator;
%  however, I see your point about using $q_i$ to connect this estimator with the
%superpopulation model.}}

@@

This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than just squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others. The variance estimation is based on a linearization
that shows that minimizing the generalized loss function subject to the
calibration constraints is asymptotically equivalent to a regression estimator.

%While this generalization is useful to an analyst who may want different
%properties of their estimator from maximum entropy estimation rather than the
%minimal squared loss, 
%% I am not sure if this is correct -----
%unless $\pi_{2i|1}^{-1} = x_i'a$ for some $a$, the
%estimator is not design consistent. 
%\footnote{ \textcolor{blue}{I do not understand this part.} 
%  \textcolor{red}{I changed the
%notation so that it would not be the unused column space notation. This is from
%Equation (11.17) of \cite{kim2024statistics}.}} 
%\footnote{\textcolor{red}{I think that I am confused about the purpose of having
%the assumption from Equation (11.17) in your sampling book. Could we discuss
%this?}}
%% --------------------------------------------

Furthermore, the regression estimator
has a nice feature that its two terms can be thought about as minimizing the
variance and bias correction,

$$ \hat Y_{reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{x_i \hat \beta_q}{\pi_{1i}}}_{
  \text{ Minimizing the variance}} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - x_i \hat \beta_q)}_{
\text{Bias correction}}.$$

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
separate have bias calibration separate from the minimizing the variance so that
we can control each in isolation. In
\cite{kwon2024debiased}, the authors show that for a generalized entropy
function $G(w)$, including a term of $g(\pi_{2i|1}^{-1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

The method of \cite{kwon2024debiased} requires known finite population 
calibration levels. It does not handle the
two-phase setup where we need to estimate the finite population total of $x$
from the Phase 1 sample. In the rest of the section, we extend this method to 
two phase sampling so that we have a valid 
estimator when including estimated Phase 1 weights with appropriate variance
estimation.

\subsection{Methodology}

@@

\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{x_i w_{2i}q_i}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{x_iq_i}{\pi_{1i}}
\end{equation}

@@

The first constraint is the existing calibration constraint and the second
ensures that design consistency is achieved. Here, 
$g(w) = \partial G / \partial w$. 
The original method of \cite{kwon2024debiased} only considered having finite
population quantities on the right hand side of \ref{eq:calconst1}.
Let $z_i = (x_i, g(\pi_{2i|1}^{-1}))$. Letting $w_{1i} = \pi_{1i}^{-1}$,
the the goal is to solve,

\begin{equation}\label{eq:primal}
  \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i = \sum_{i \in A_1} w_{1i} z_i
  q_i.
\end{equation}

Let $\hat w_{2i|1}$ be the solution to Equation~\ref{eq:primal}, then the
estimate of $Y_N$ is $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i|1} y_i$.

\subsection{Theoretical Results}

%\begin{itemize}
%  \item[(A1)] $G(w)$ is strictly convex and twice differentiable in an interval
%    $\mathcal{V} \in \R$,
%  \item[(A2)] There exists $c_1, c_2 \in \mathcal{V}$ with $c_1, c_2 > 0$ such
%    that $c_1 < \pi_{2i|1} < c_2$ for $i = 1, \dots, n_2$, and 
%    that $c_1 < \pi_{1i} < c_2$ for $i = 1, \dots, n_1$,
%  \item[(A3)] If $\pi_{2ij|1}$ is joint inclusion probability of elements $i$
%    and $j$ in $A_2$, the Phase 2 sample and 
%    $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$ then,
%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{2ij|1}| <
%    \infty,$$
%    and if $\pi_{1ij}$ is the joint inclusion probability of elements $i$ and
%    $j$ in $A_1$, the Phase 1 sample and 
%    $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ then,
%    $$\limsup_{n \to \infty} \max_{i,j \in U: i \neq j} |\Delta_{1ij}| <
%    \infty,$$
%  \item[(A4)] Assume that $\Sigma_z = \lim_{N \to \infty} \sum_{i \in U} z_i
%    z_i^T/N$ exists and is positive definite, the average fourth moment of $(y_i,
%    x_i^T)$ is finite ($\limsup_{N \to \infty} \sum_{i \in U} ||(y_i, x_i^T)||^4
%    / N < \infty$), and $\Gamma(\lambda) = \lim_{N \to \infty}  \sum_{i \in U}
%    f'(\lambda^T z_i) z_i z_i^T / N$ exists in a neighborhood around $\lambda_0
%    = (\bf 0, 1)$.
%\end{itemize}

%\begin{theorem}[Design Consistency]
%  Suppose that Conditions (A1) - (A4) hold. Then the solution $\hat w$ to 
%  Equation~\ref{eq:primalloss} subject to Equation~\ref{eq:calconst1} and 
%  Equation~\ref{eq:calconst2} 
%  exists and is unique with probability approaching 1. Furthermore, the
%  estimator $\hat \theta_{DCE} = \sum_{i \in A_2} \frac{\hat w_i y_i}{\pi_{1i}}$
%  satisfies
%
%  $$\hat \theta_{DCE} = \hat \theta_{DC} + o_p(n_2^{-1})$$
%
%  where $\hat \theta_{DC}$ is the debiased calibration estimator of 
%  \cite{kwon2024debiased}.
%
%\end{theorem}
%
%\begin{proof}
%  This proof follows the proof of \cite{kwon2024debiased} in a straightforward
%  manner. There are two main modifications that need to be made. The first is
%  that the proof needs to be written for two-phase sampling with $n_1$ and $n_2$
%  instead of a single $n$. The second is that we need to ensure that the result
%  still holds with an estimated population constraint that is $O_p(n_1^{-1/2})$
%  consistent.
%
%  With the Conditions (A1) - (A4) written with respect the the two-phase sample,
%  the first challenge is done. The second step largely holds from the existing
%  proof of \cite{kwon2024debiased} with modifications such as 
%
%  $$\hat Q(\lambda) = n_2^{-1} \sum_{i \in A_2} F(\lambda^T z_i^*) - n_1^{-1}
%  \lambda^T \sum_{i \in A_1} z_i^*$$ 
%
%  ensuring that our extension holds.
%
%%  \textcolor{red}{Dr. Kim, I know that I need to write up this proof in more
%%  detail, but I am unsure about the level of granularity that is required now. I
%%  can basically rewrite everything from the Appendix of \cite{kwon2024debiased}
%%  with the two-phase modification to show this result, but I don't know if I
%%  should yet.}
%\end{proof}

%We need the following additional assumptions to prove asymptotic normality.
%
%\begin{itemize}
%  \item[(B1)] 
%  \item[(B2)] The HT estimator is asymptotically normal under the sampling
%    design in the sense that 
%
%\end{itemize}

\begin{theorem}[Variance Estimation]
  Let $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$ and 
  $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
  - z_i \hat \phi / q_i\right)$ where

  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$

  An unbiased estimator of the variance of $\hat Y_{DCE}$ is

  \begin{align*}
    \hat V 
    &= \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
    \pi_{1i}\pi_{1j}}{\pi_{1ij}}
    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right) \\ 
    &+ \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
    \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
  \end{align*}
\end{theorem}

\begin{proof}
  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
  asymptotic variance of this estimator using the linearization technique of
  \cite{randles1982asymptotic}.

  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
  Lagrange multipliers we have a Lagrangian of

  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$

  Our first order conditions yield,

  \begin{align*}
    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
  \end{align*}

  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
  \hat w_{2i}(\hat \lambda) y_i$.

  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
  \cite{randles1982asymptotic}. First, notice that

  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
  \sum_{i \in A_1} w_{1i} z_i -
  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
  := \hat Y_\ell(\hat \lambda, \phi).$$

  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
  is the probability limit of $\hat \lambda$. Notice, that 
  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
  that the optimal value of $\phi$ solves,

  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
    w_{2i|1}(\lambda) / q_i \right]$$

  which means that

  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$

  Hence,

  \begin{align*}
    \hat Y_\ell(\lambda^*, \phi^*) 
    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
    &= \sum_{i \in A_1} w_{1i} \eta_i
  \end{align*}

  where

  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
  - z_i \phi^* / q_i\right).$$

  To analyze the variance, consider the reverse framework of
  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
  known at the
  finite population level. Then with this we have,

  \begin{align*}
    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
  \end{align*}

  We can estimate this variance using

  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
  $$

  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
  - z_i \hat \phi / q_i\right)$ where

  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$

  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
  asymptotically equivalent to
  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of

  \begin{align*}
    \text{Bias}(\hat V_{DCE}) 
    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
  \end{align*}

  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
  However, we can estimate this bias with

  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$

\end{proof}

\subsection{Simulation Studies}

@@

\section{Topic 2: Non-nested Two-Phase Sampling}

@@

Now we consider the sampling mechanism known as non-nested two phase sampling 
(\cite{kim2024statistics}). In the last section, we considered two phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two phase sampling the Phase 2 sample is independent of the Phase 1 sample. It
is a separate independent sample of the same population. Like traditional two
phase sampling, we consider the Phase 1 sample, $A_1$, to consist of
observations of $(X_i)_{i = 1}^{n_1}$ and the Phase 2 sample, $A_2$, to consist
of observations of $(X_i, Y_i)_{i = 1}^{n_2}$.

Whereas the classical two phase estimator uses a single Horvitz-Thompson
estimator of the Phase 1 sample to construct estimates for calibration totals,
in the non-nested two phase sample we have two independent Horvitz-Thompson
estimators of the total of $X$,

$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$
